---
title: 'Not Always A Matter Of Context: Preliminary data summaries'
author: "Nicole LaGrange, Tania Delgado, Pamela Riviere"
date: '2017-12-14'
output:
  html_document: default
  pdf_document: default
---

####### __Paper.__ Buechner and Maier (2016) Not always a matter of context: direct effects of red on arousal but context-dependent moderations on valence. _PeerJ_

####### __Summary.__ This study aimed to determine whether colors intrinsically elicit different degrees of physiological arousal, or whether their perceptual effects are dictated by the contextual circumstances in which the colors are experienced. To investigate this question, Buechner and Maier (2016) sorted subjects into one of ___four experimental categories___. These categories consisted of assessing the arousal and valence properties of either:

* four "test-relevant" pictures framed in red
* four "test-relevant" pictures framed in blue
* four "neutral" pictures framed in red
* four "neutral" pictures framed in blue

To assess arousal, participants in each category were asked: "Describe how physiologically arousing do you think [the situation in this picture] is". To assess valence, participants in each category were asked: "Describe how unpleasant do you think [the situation in this picture] is". In response to each question, participants assigned each picture a number from 1 (not at all) to 9 (extremely). 

######## The dataframe contains rows of subjects with their corresponding assessement of a given picture as physically arousing or unpleasant (a measure of valence).

```{r, warning=FALSE, message=FALSE}
library(foreign)
library(dplyr)
library(tidyverse)

color_data_pr = read.csv("color_data_pr.csv")
```

### Replicating Analyses and Figures

######### TODO: Remember to include commentary as to why it's potentially weird to collapse across judgments for different pictures, include error bars, run anova stats, polish the names of axis labels, Scheffe post hoc

######### TODO: Remember to 
* include commentary as to why it's potentially weird to collapse across judgments for different pictures
* include error bars
* run anova stats
* polish the names of axis labels
* Scheffe post hoc


```{r}
ggplot(color_data_pr, aes(x=Condition, Pic_valence_MEAN)) +
  geom_bar(stat='summary', fun.y='mean') +
  stat_summary(fun.y = function(x){mean(x)},
               fun.ymin = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
               fun.ymax = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
               geom= 'pointrange') +
  labs(title = 'Mean Picture Valence Across Conditions', x = "Condition", y = "Mean Picture Valence") +
  theme_minimal()

ggplot(color_data_pr, aes(x=pic, Pic_arousal_MEAN, fill=color)) +
  geom_bar(stat="summary", fun.y="mean", position="dodge") + 
  labs(x = "Condition", y = "Mean Picture Valence Rating", title = 'Mean Picture Valence Across Conditions') +
  theme_minimal()

ggplot(color_data_pr, aes(x=Condition, Pic_arousal_MEAN)) +
  geom_bar(stat="summary", fun.y="mean", position="dodge") +
  stat_summary(fun.y = function(x){mean(x)},
               fun.ymin = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
               fun.ymax = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
               geom= 'pointrange') +
  labs(x = "Condition", y = "Mean Picture Arousal Rating", title = 'Mean Picture Arousal Across Conditions') + 
  theme_minimal()

ggplot(color_data_pr, aes(x=pic, Pic_arousal_MEAN, fill=color)) +
  geom_bar(stat="summary", fun.y="mean", position="dodge", width = 0.5) + 
  stat_summary(fun.y = function(x){mean(x)},
               fun.ymin = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
               fun.ymax = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
               geom= 'pointrange', position = position_dodge(0.5)) +
  labs(title = 'Mean Arousal Rating for Neutral v. Test Pictures', y = 'Mean Picture Arousal Rating', x = 'Picture Type') + 
  theme_minimal()

summary(aov(data=color_data_pr, Pic_arousal_MEAN ~ color*pic))

ggplot(color_data_pr, aes(x=color, Pic_arousal_MEAN)) +
  geom_bar(stat="summary", fun.y="mean", position="dodge") + 
  labs(title = 'Mean Arousal Rating for Blue v. Red Pictures', y = 'Mean Picture Arousal Rating', x = 'Picture Color') + theme_minimal()
```

```{r}
ggplot(data = color_data_pr, aes(x=color, y=Pic_arousal_MEAN)) +
geom_bar(stat="summary", fun.y="mean", position="dodge") +
  stat_summary(fun.y = function(x){mean(x)},
               fun.ymin = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
               fun.ymax = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
               geom= 'pointrange') +
  labs(title = 'Mean Arousal Rating for Blue v. Red Pictures', y = 'Mean Picture Arousal Rating', x = 'Picture Color') + theme_minimal()

summary(aov(data = color_data_pr, Pic_arousal_MEAN ~ color))
```
  
```{r}
### Beyond the paper
########## TODO: Run ANCOVA, using trait and test anxiety as the controls, figures we chose to improve, why are they improvements (violin plots here and why they are a nicer representation of the distribution)

### Violin Plot for Arousal Mean
ggplot(color_data_pr, aes(x=pic, Pic_arousal_MEAN, fill=color)) + geom_violin() +   labs(title = 'Mean Arousal Rating for Blue v. Red Pictures', y = 'Mean Picture Arousal Rating', x = 'Picture Type') + theme_minimal()

### Violin Plot for Valence Mean
ggplot(color_data_pr, aes(x=pic, Pic_valence_MEAN, fill=color)) + geom_violin() +   labs(title = 'Mean Valence Rating for Blue v. Red Pictures', y = 'Mean Picture Valence Rating', x = 'Picture Type') + theme_minimal()

# bar or violin plots for the distribution of arousal and valence judgements for each picture
arousal_by_pic = c(mean(color_data_pr$pic1_arousal), 
                   mean(color_data_pr$pic2_arousal), 
                   mean(color_data_pr$pic3_arousal), 
                   mean(color_data_pr$pic4_arousal))

valence_by_pic = c(mean(color_data_pr$pic1_valence), mean(color_data_pr$pic2_valence), mean(color_data_pr$pic3_valence), mean(color_data_pr$pic4_valence))

pics = c('pic1', 'pic2', 'pic3', 'pic4','pic1', 'pic2', 'pic3', 'pic4')
new.df = data.frame(arousal_by_pic, valence_by_pic, pics)

labels = c('arousal', 'arousal', 'arousal', 'arousal',
           'valence', 'valence', 'valence', 'valence')

ratings = c(arousal_by_pic, valence_by_pic)
new.df2 = data.frame(ratings, pics, labels)

ggplot(new.df2, aes(x=pics, ratings, fill=labels)) +
  geom_bar(stat="identity", position="dodge") + 
  labs(title = 'Mean Picture Arousal Across Pictures') + 
  theme_minimal()

# try this for the error bars
stat_summary (fun.y = function(x){mean(x)},
                fun.ymin = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.ymax = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange')
```

__some ANOVAs.__ 
``` {r some anovas}

# Let's look at main and interaction effects
interact.val = anova(lm(data = color_data_pr, Pic_valence_MEAN ~ color*pic))
interact.aro = anova(lm(data = color_data_pr, Pic_arousal_MEAN ~ color*pic))


#is this a valid way to go about this? - tania
anova.val.cond = anova(lm(data = color_data_pr, Pic_valence_MEAN ~ Condition))
anova.aro.cond = anova(lm(data = color_data_pr, Pic_arousal_MEAN ~ Condition))

interact.val.cond = anova(lm(data = color_data_pr, Pic_valence_MEAN ~ Condition*Pic_arousal_MEAN))
interact.aro.cond = anova(lm(data = color_data_pr, Pic_arousal_MEAN ~ Condition*Pic_valence_MEAN))

```

__some ANCOVAs.__
``` {r some ancovas}

#is this the right way to go about this? - tania

# ANCOVA with valence rating as the response variable, trait or test anxiety as controls
ancova.val.trait_anxiety = anova(lm(data = color_data_pr, Pic_valence_MEAN ~ STAI_MEAN + Condition))
ancova.val.test_anxiety = anova(lm(data = color_data_pr, Pic_valence_MEAN ~ AEQanxiety_MEAN + Condition))

# ANCOVA with arousal rating as the response variable, trait or test anxiety as controls
ancova.aro.trait_anxiety = anova(lm(data = color_data_pr, Pic_arousal_MEAN ~ STAI_MEAN + Condition))
ancova.aro.test_anxiety = anova(lm(data = color_data_pr, Pic_arousal_MEAN ~ AEQanxiety_MEAN + Condition))

```

__Tania messing around to understand the data/feel free to ignore.__
```{r Tania messing around with things to understand the data better, enter at your own risk}

## First let's create a simpler dataframe to work with

simple.color.df <- data.frame(condition=color_data_pr$Condition,pic1_valence=color_data_pr$pic1_valence,pic2_valence=color_data_pr$pic2_valence,pic3_valence=color_data_pr$pic3_valence,pic4_valence=color_data_pr$pic4_valence,pic1_arousal=color_data_pr$pic1_arousal,pic2_arousal=color_data_pr$pic2_arousal,pic3_arousal=color_data_pr$pic3_arousal,pic4_arousal=color_data_pr$pic4_arousal,test.anxiety=color_data_pr$STAI_MEAN,trait.anxiety=color_data_pr$AEQanxiety_MEAN)

## Let's look at the data by image and condition

# Red and Neutral condition
red.neu = filter(simple.color.df, condition=='red_neu')

# Red/neutral valence ratings
means.val.red.neu = c(mean(red.neu$pic1_valence),
                     mean(red.neu$pic2_valence),
                     mean(red.neu$pic3_valence),
                     mean(red.neu$pic4_valence))

pic1.val.mean.rn = means.val.red.neu[1]
pic2.val.mean.rn = means.val.red.neu[2]
pic3.val.mean.rn = means.val.red.neu[3]
pic4.val.mean.rn = means.val.red.neu[4]

val.red.neu.avg = mean(c(mean(red.neu$pic1_valence),
                     mean(red.neu$pic2_valence),
                     mean(red.neu$pic3_valence),
                     mean(red.neu$pic4_valence)))

# Red/neutral arousal ratings
means.aro.red.neu = c(mean(red.neu$pic1_arousal),
                     mean(red.neu$pic2_arousal),
                     mean(red.neu$pic3_arousal),
                     mean(red.neu$pic4_arousal))

pic1.aro.mean.rn = means.aro.red.neu[1]
pic2.aro.mean.rn = means.aro.red.neu[2]
pic3.aro.mean.rn = means.aro.red.neu[3]
pic4.aro.mean.rn = means.aro.red.neu[4]

aro.red.neu.avg = mean(c(mean(red.neu$pic1_arousal),
                     mean(red.neu$pic2_arousal),
                     mean(red.neu$pic3_arousal),
                     mean(red.neu$pic4_arousal)))

# Red/test condition

red.test = filter(simple.color.df, condition=='red_test')

# Red/test valence ratings
means.val.red.test = c(mean(red.test$pic1_valence),
                     mean(red.test$pic2_valence),
                     mean(red.test$pic3_valence),
                     mean(red.test$pic4_valence))

pic1.val.mean.rt = means.val.red.test[1]
pic2.val.mean.rt = means.val.red.test[2]
pic3.val.mean.rt = means.val.red.test[3]
pic4.val.mean.rt = means.val.red.test[4]

val.red.test.avg = mean(c(mean(red.test$pic1_valence),
                     mean(red.test$pic2_valence),
                     mean(red.test$pic3_valence),
                     mean(red.test$pic4_valence)))

# Red/test arousal ratings
means.aro.red.test = c(mean(red.test$pic1_arousal),
                     mean(red.test$pic2_arousal),
                     mean(red.test$pic3_arousal),
                     mean(red.test$pic4_arousal))

pic1.aro.mean.rt = means.aro.red.test[1]
pic2.aro.mean.rt = means.aro.red.test[2]
pic3.aro.mean.rt = means.aro.red.test[3]
pic4.aro.mean.rt = means.aro.red.test[4]

aro.red.test.avg = mean(c(mean(red.test$pic1_arousal),
                     mean(red.test$pic2_arousal),
                     mean(red.test$pic3_arousal),
                     mean(red.test$pic4_arousal)))

# Blue/neutral condition
blue.neu = filter(simple.color.df, condition=='blue_neu')

# blue/neutral valence ratings


```


